---
title: "Support Vector Machines"
output: html_notebook
---

# Intro to Support Vector Machines (SVMs)

[Intro to SVMs](https://www.datacamp.com/tutorial/support-vector-machines-r)

### Create some sample data

```{r}
set.seed(10111)
# Make a matrix x, normally distributed with 20 observations in 2 classes/variables.
x <- matrix(rnorm(40), 20, 2)
x
# Make a y variable which is going to be either -1 or 1, with 10 in each class.
y <- rep(c(-1, 1), c(10, 10))
y
# For y = 1, move the means from 0 to 1 in each of the coordinates.
x[y == 1, ] <- x[y == 1, ] + 1 # Alternative for loop notation to increase values in rows 11 to 20 by 1.
x
plot(x, col = y + 3, pch = 19) # pch = 19 creates big visible dots coded blue or red according to whether the response is 1 or -1.
```

### Creating the model

```{r}
# Load the package `e1071` which contains the `svm` function.
library(e1071)
# Turn the data into a data frame and make `y` a factor variable.
dat <- data.frame(x, y = as.factor(y))
dat
# Train the model on the data
svmfit <- svm(y ~ ., data = dat, kernel = 'linear', cost = 10, scale = F)
print(svmfit)
```

"No. of support vectors: 6" are the points that are close tot he boundary or on the wrong side of the boundary.

```{r}
# Plot of SVM that shows the decision boundary.
plot(svmfit, dat)
```

### Make your own plot + Make predictions

The first thing to do is to create a grid of values or a lattice of values for x1 and x2 that covers the whole domain on a fairly fine lattice. To do so, create a function called `make.grid()`. It takes in your data matrix, as well as an argument n which is the number of points in each direction. Here you're going to ask for a 75 x 75 grid.

```{r}
make.grid <- function(data_matrix, size = 75) {
  # Get the range of each of the variables in the data_matrix
  grid_range = apply(data_matrix, 2, range)
  # print(grid_range)
  # Use the seq function to go from the lowest to the upper value to make a grid of length = size
  x1 = seq(from = grid_range[1, 1], to = grid_range[2, 1], length = size)
  x2 = seq(from = grid_range[1, 2], to = grid_range[2, 2], length = size)
  # Make the lattice using expand.grid()
  expand.grid(X1 = x1, X2 = x2)
}
```

Apply `make.grid()` on x and see the first few values of the lattice from 1 to 10.

```{r}
xgrid <- make.grid(x)
xgrid[1:10, ]
```

As you can see, the grid goes through the 1st coordinate first, holding the 2nd coordinate fixed.

Having made the lattice, make a prediction at each point in the lattice.

```{r}
# Use predict() and name the response 'ygrid'.
ygrid <- predict(svmfit, xgrid)
ygrid
# Plot and colour code the points according to the classification so that the decision boundary is clear.
plot(xgrid, col = c("red", "blue")[as.numeric(ygrid)], pch = 20, cex = 0.2)
# Put the original points on the plot.
points(x, col = y + 3, pch = 19)
# scmfit$index tells which are the support points. You can include them in the plot.
points(x[svmfit$index, ], pch = 5, cex = 2)
```

Extract the linear coefficients for the decision boundary.

```{r}
beta <- drop(t(svmfit$coefs) %*% x[svmfit$index, ])
beta0 <- svmfit$rho
```

With these coefficients, you can draw the decision boundary using the equation of the form:

$$\beta_0 + \beta_1 + x1 + \beta_2 * x2 = 0$$

From this equation, figure out a slop and an intercept for the decision boundary. Use the function `abline` with those 2 arguments. The subsequent 2 `abline` functions represent the upper and lower margins of the decision boundary, respectively.

```{r}
# Same as previously
plot(xgrid, col = c("red", "blue")[as.numeric(ygrid)], pch = 20, cex = .2)
points(x, col = y + 3, pch = 19)
points(x[svmfit$index,], pch = 5, cex = 2)
# Add the decision boundary
abline(beta0 / beta[2], -beta[1] / beta[2])
abline((beta0 - 1) / beta[2], -beta[1] / beta[2], lty = 2)
abline((beta0 + 1) / beta[2], -beta[1] / beta[2], lty = 2)
```

